---
title: "Binomial analysis analysis"
author: "Bodo"
date: "16/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse) # for data processing
library(brms) # for bayesian models
library(effsize) # for cohen's d
library(patchwork) # for double plots
library(tidybayes) # for half-eye plot
```

Load data:

```{r, message = FALSE, warning = FALSE}
binoms <- read_csv('../data/binomial_all_frequencies.csv')
```

Quick glance at the structure:

```{r}
binoms
```

The pairs are:

```{r}
pairs <- tibble(pair = 1:7,
                name = c('add and/or subtract', 'plus and/or minus',
                         'addition and/or subtraction', 'more and/or less',
                         'increase and/or decrease', 'many and/or few',
                         'most and/or least'))
```

In this analysis, we are not interested in the difference for and/or, so we'll average over that:

```{r}
binoms <- binoms %>% 
  group_by(pair, type, file, register) %>% 
  summarize(freq = sum(freq))
```

Show:

```{r}
binoms
```

## Descriptive analysis

Calculate averages:

```{r}
binoms %>% 
  group_by(type) %>% 
  summarize(total = sum(freq))
```

Calculate averages per pair:

```{r}
binom_pairs <- binoms %>% 
  group_by(pair, type) %>% 
  summarize(total = sum(freq))

# Show:

binom_pairs
```

Calculate log frequencies:

```{r}
binom_pairs <- mutate(binom_pairs,
                      log_freq = log10(total + 1))
```

Calculate ratios:

```{r}
add_first_totals <- binom_pairs[rep(c(TRUE, FALSE), 7), ]$total
add_second_totals <- binom_pairs[rep(c(FALSE, TRUE), 7), ]$total

ratios <- add_first_totals / add_second_totals
ratios <- tibble(pair = binom_pairs[rep(c(TRUE, FALSE), 7), ]$pair,
       ratio = ratios)

left_join(pairs, ratios)
```

Calculate Cohen's d:

```{r}
cohen.d(log_freq ~ type | Subject(pair), data = binom_pairs, paired = TRUE)
```

Get the labels for annotation:

```{r}
pairs$annotate_freqs <- filter(binom_pairs, type == 'sub_first')$log_freq

pairs
```

Make a plot of this:

```{r}
# Define plot and aesthetics:

binom_p <- binom_pairs %>% 
  mutate(type = ifelse(type == 'add_first',
                       'addition\nfirst', 'subtraction\nfirst')) %>% 
  ggplot(aes(x = type, y = log_freq, fill = type, group = pair))

# Add geoms:

binom_p <- binom_p +
  geom_line(col = 'grey') +
  geom_point(size = 3, shape = 21, alpha = 0.85) +
  annotate(geom = 'text',
           x = rep(2.1, nrow(pairs)),
           y = pairs$annotate_freqs,
           label = pairs$name,
           hjust = 0,
           col = 'grey33')

# Add scales and axes labels:

binom_p <- binom_p +
  scale_y_continuous(breaks = seq(0, 4, 1)) +
  scale_fill_manual(values = c("#E69F00", "#0072B2")) +
  coord_cartesian(xlim = c(1.3, 2.3), ylim = c(0, 4),
                  clip = 'off') +
  xlab(NULL) +
  ylab('Average log10 frequency')

# Add themes:

binom_p <- binom_p +
  theme_classic() +
  theme(legend.position = 'none') +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 15,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.text.x = element_text(face = 'bold', size = 10),
        plot.margin = margin(r = 75, l = 5, b = 5, t = 5))

# Show and save:

binom_p
ggsave(plot = binom_p,
       filename = '../figures/pdf/binomial_frequencies.pdf',
       width = 4, height = 4)
ggsave(plot = binom_p,
       filename = '../figures/png/binomial_frequencies.png',
       width = 4, height = 4)
ggsave(plot = binom_p,
       filename = '../figures/tiff/binomial_frequencies.tiff',
       width = 4, height = 4)
```

## Bayesian model

How many are zeros?

```{r}
sum(binoms$freq == 0) / nrow(binoms)
```

Look at the number of zeros as a function of add/subtract first:

```{r}
binoms %>% 
  mutate(is_zero = freq == 0) %>%
  group_by(type) %>% 
  summarize(zeros = sum(is_zero)) %>% 
  mutate(prop = zeros / sum(zeros),
         prop = round(prop, 2))
```

Make `sub_first` the reference level:

```{r}
binoms <- mutate(binoms,
                 type = factor(type,
                               levels = c('sub_first', 'add_first')))
```

Weakly informative prior:

```{r}
weak_prior <- prior(normal(0, 2), class = 'b')
```

Model this with a negative binomial regression model:

```{r}
binom_mdl <- brm(bf(freq ~ 1 + type +
                      (1 + type|pair) + (1 + type|file) + (1 + type|register),
                    zi ~ 1),
                 data = binoms,
                 
                 prior = weak_prior,
               
                 family = zero_inflated_negbinomial('log'),
               
                 # MCMC settings:
                 
                 seed = 666, chains = 4, cores = 4,
                 warmup = 4000, iter = 6000,
                 control = list(adapt_delta = 0.99))
```

Save the model:

```{r}
save(binom_mdl, file = '../models/binomial_model.RData')
```

Check the model:

```{r}
binom_mdl
```

Perform posterior predictive checks (with empirical cumulative distribution function):

```{r}
pp_check(binom_mdl,
         ndraws = 100,
         type = 'ecdf_overlay')
```

Perform a test of the posterior probability of the effect being above/below zero.

```{r}
hypothesis(binom_mdl, 'typeadd_first < 0')
hypothesis(binom_mdl, 'typeadd_first > 0')
```

Do this by hand by extracting posterior samples:

```{r}
binom_samples <- posterior_samples(binom_mdl)
```

Plot the posterior:

```{r}
post_p <- binom_samples %>% 
  ggplot(aes(x = b_typeadd_first)) +
  stat_halfeye(fill = 'steelblue', alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = 'dashed')

post_p <- post_p +
  xlab('Addition coefficient') +
  ylab('Probability density') +
  coord_cartesian(y = c(0, 1.1), clip = 'off') +
  scale_y_continuous(breaks = seq(0, 1, 0.25))

post_p <- post_p +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(face = 'bold', size = 14,
                                    margin = margin(t = 10)),
        axis.title.y = element_text(face = 'bold', size = 16,
                                    margin = margin(r = 12)))

post_p
```

Put the raw frequencies and the posterior into a double plot:

```{r}
# Tweak plots for double plotting:

binom_p <- binom_p +
  ggtitle('(a) Log frequencies') +
  theme(plot.title = element_text(face = 'bold'),
        plot.title.position = 'plot')

post_p <- post_p +
  ggtitle('(b) Posterior distribution of frequency difference') +
  theme(plot.title = element_text(face = 'bold'),
        plot.margin = margin(r = 25),
        plot.title.position = 'plot')

# Merge:

both_p <- binom_p + post_p

# Show and save:

both_p
ggsave(filename = '../figures/pdf/binomial_double_plot.pdf',
       width = 10, height = 4.5)
ggsave(filename = '../figures/png/binomial_double_plot.png',
       width = 10, height = 4.5)
ggsave(filename = '../figures/tiff/binomial_double_plot.tiff',
       width = 10, height = 4.5)
```


