---
title: "Linking semantics with frequency and valence"
author: "Bodo"
date: "16/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse) # for data processing
library(lsa) # for cosine similarity function
library(brms) # for bayesian models
library(effsize) # for cohen's d
library(patchwork) # for double plots
library(tidybayes) # for halfeye plots
```

Load data:

```{r, message = FALSE, warning = FALSE}
verbs <- read_csv('../data/verb_cosines.csv')

# raw frequencies

COCA <- read_csv('../data/COCA_freq_CD.csv')
war <- read_csv('../data/warriner_2013_valence_arousal.csv')
```

Preprocess COCA:

```{r}
COCA <- select(COCA, w1,
               freq_coca_spok:freq_coca_acad)
```

Merge COCA freqs into verbs data frame:

```{r}
verbs <- left_join(verbs, COCA, by = c('word' = 'w1'))
```

Get cumulative frequencies (across registers) for simple reporting:

```{r}
verbs <- mutate(verbs,
                freq = freq_coca_spok +
                  freq_coca_fic +
                  freq_coca_mag +
                  freq_coca_news +
                  freq_coca_acad)
```

Join the Warriner et al. (2013) valence in here:

```{r}
verbs <- left_join(verbs, select(war, Word, V.Mean.Sum),
                   by = c('word' = 'Word'))
```

Compute log frequency:

```{r}
verbs <- mutate(verbs,
                log_freq = log10(freq))
```

## Descriptive statistics

Compute correlation with frequency:

```{r}
with(verbs, cor(log_freq, diffs, use = 'complete.obs'))
```

Compute correlation with semantic prosody:

```{r}
with(verbs, cor(V.Mean.Sum, diffs, use = 'complete.obs'))
```

## Bayesian models

For regression analysis standardize both:

```{r}
verbs <- mutate(verbs,
                # centering:
                
                diff_z = diffs - mean(diffs),
                log_freq_z = log_freq - mean(log_freq, na.rm = TRUE),
                val_z = V.Mean.Sum - mean(V.Mean.Sum, na.rm = TRUE),
                
                # standardizing:
                
                diff_z = diff_z / sd(diff_z),
                log_freq_z = log_freq_z / sd(log_freq_z, na.rm = TRUE),
                val_z = val_z / sd(val_z, na.rm = TRUE))
```


Regress addition bias onto frequency and context valence:

```{r}
verb_mdl <- brm(diff_z ~ 1 + log_freq_z + val_z,
                data = verbs,

                # likelihood:
                
                family = gaussian,                
                                
                # prior:
                
                prior = prior(normal(0, 1), class = 'b'),
                
                seed = 666, chains = 4, cores = 4,
                warmup = 2000, iter = 4000)
```

Save:

```{r}
save(verb_mdl, file = '../models/verb_mdl.RData')
```

Show the results:

```{r}
verb_mdl
```

Perform hypothesis tests for frequency:

```{r}
hypothesis(verb_mdl, 'log_freq_z < 0')
hypothesis(verb_mdl, 'log_freq_z > 0')
```

Same for valence:

```{r}
hypothesis(verb_mdl, 'val_z < 0')
hypothesis(verb_mdl, 'val_z > 0')
```

## Frequency plot

Plot marginal effect of frequency:

```{r}
conditional_effects(verb_mdl, 'log_freq_z')
```

Superimpose the predictions on a scatterplot:

```{r}
# Plot basics:

freq_p <- verbs %>%
  ggplot(aes(x = log_freq_z, y = diff_z)) +
  geom_point(alpha = 0.5, fill = 'steelblue',
             pch = 21) +
  geom_abline(mapping = aes(intercept = fixef(verb_mdl)[1, 1],
                            slope = fixef(verb_mdl)[2, 1]),
              size = 1.2)

# Axes and labels:

freq_p <- freq_p +
  xlab('Log frequency (z-scored)') +
  ylab('Addition bias (z-scored)')

# Add cosmetics:

freq_p <- freq_p +
  theme_classic() +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 12,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.title.x = element_text(margin = margin(t = 10, r = 0,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.text.x = element_text(face = 'bold', size = 10),
        axis.text.y = element_text(face = 'bold', size = 10))

# Show and save:

freq_p
ggsave(plot = freq_p,
       filename = '../figures/pdf/frequency_bias.pdf',
       width = 5.5, height = 4)
ggsave(plot = freq_p,
       filename = '../figures/png/frequency_bias.png',
       width = 5.5, height = 4)
ggsave(plot = freq_p,
       filename = '../figures/tiff/frequency_bias.tiff',
       width = 5.5, height = 4)
```

Plot marginal effect of valence:

```{r}
conditional_effects(verb_mdl, 'val_z')
```

Superimpose the predictions on a scatterplot:

```{r}
# Plot basics:

val_p <- verbs %>%
  ggplot(aes(x = val_z, y = diff_z)) +
  geom_point(alpha = 0.5, fill = 'steelblue',
             pch = 21) +
  geom_abline(mapping = aes(intercept = fixef(verb_mdl)[1, 1],
                            slope = fixef(verb_mdl)[3, 1]),
              size = 1.2)

# Axes and labels:

val_p <- val_p +
  xlab('Valence (z-scored)') +
  ylab('Addition bias (z-scored)')

# Add cosmetics:

val_p <- val_p +
  theme_classic() +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 12,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.title.x = element_text(margin = margin(t = 10, r = 0,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.text.x = element_text(face = 'bold', size = 10),
        axis.text.y = element_text(face = 'bold', size = 10))

# Show and save:

val_p
ggsave(plot = val_p,
       filename = '../figures/pdf/valence_bias.pdf',
       width = 5.5, height = 4)
ggsave(plot = val_p,
       filename = '../figures/png/valence_bias.png',
       width = 5.5, height = 4)
ggsave(plot = val_p,
       filename = '../figures/tiff/valence_bias.tiff',
       width = 5.5, height = 4)
```

Put them both together:

Put the raw frequencies and the posterior into a double plot:

```{r}
# Tweak plots for double plotting:

freq_p <- freq_p +
  ggtitle('(a) Addition bias ~ frequency') +
  theme(plot.title = element_text(face = 'bold'),
        plot.title.position = 'plot')

val_p <- val_p +
  ggtitle('(b) Addition bias ~ valence') +
  theme(plot.title = element_text(face = 'bold'),
        plot.title.position = 'plot') +
  ylab(NULL)

# Merge:

both_p <- freq_p + val_p

# Show and save:

both_p
ggsave(plot = both_p,
       filename = '../figures/pdf/frequency_valence_double_plot.pdf',
       width = 11, height = 4.5)
ggsave(plot = both_p, 
       filename = '../figures/png/frequency_valence_double_plot.png',
       width = 11, height = 4.5)
ggsave(plot = both_p, 
       filename = '../figures/tiff/frequency_valence_double_plot.tiff',
       width = 11, height = 4.5)
```

Show the posteriors:

```{r}
# Basic plot:

post_p <- verb_mdl %>% 
  spread_draws(b_log_freq_z,
               b_val_z) %>% 
  pivot_longer(cols = b_log_freq_z:b_val_z,
               names_to = 'coefficient',
               values_to = 'estimate') %>% 
  mutate(coefficient = ifelse(coefficient == 'b_val_z',
                              'valence', 'frequency')) %>% 
  ggplot(aes(x = estimate, y = coefficient)) +
  stat_halfeye(fill = 'steelblue') +
  geom_vline(xintercept = 0, linetype = 'dashed')

# Axes and labels:

post_p <- post_p +
  coord_cartesian(xlim = c(-0.2, 0.7)) +
  scale_x_continuous(breaks = seq(-0.2, 0.6, 0.2)) +
  xlab('Posterior estimate') +
  ylab(NULL)

# Cosmetics:

post_p <- post_p +
  theme_classic() +
  theme(axis.text.y = element_text(size = 14,
                                   face = 'bold'),
        axis.title.x = element_text(margin = margin(t = 10, r = 0,
                                                    b = 0, l = 0),
                                    size = 14, face = 'bold'),
        axis.text.x = element_text(face = 'bold', size = 10))

# Show and save:

post_p
ggsave(plot = post_p,
       filename = '../figures/pdf/valence_frequency_posteriors.pdf',
       width = 6.5, height = 4)
ggsave(plot = post_p,
       filename = '../figures/png/valence_frequency_posteriors.png',
       width = 6.5, height = 4)
ggsave(plot = post_p,
       filename = '../figures/tiff/valence_frequency_posteriors.tiff',
       width = 6.5, height = 4)
```










